---
title: "MarineGEO data tracker development"
output: html_notebook
---

Development of the graphs for the MarineGEO data tracker dashboard

```{r include=FALSE}
library(tidyverse)
library(wordcloud)
library(tm)
library(here)
library(forcats)
set_here(path="../")
```


# Load the csv file into a data frame

```{r}
tracker <- read_csv(here("data/data-status-tracker.csv"))
tracker
```


# Quick check - frequency of projects
```{r}
tracker %>% group_by(Project) %>% count()

```


## Word Cloud of all the project names

Word cloud based on frequency. Project names that are more frequent are larger. Note: this splits project by word (ie Seagrass Monitoring is "Seagrass" & "Monitoring").

```{r, fig.width=5, fig.height=5}
d <- Corpus(VectorSource(tracker$Project)) %>% tm_map(content_transformer(tolower))

#inspect(d)

tdm <- TermDocumentMatrix(d) %>%
  as.matrix()

v <- sort(rowSums(tdm), decreasing = TRUE)
d <- data.frame(word = names(v), freq=v)

set.seed(12)
wordcloud(words=d$word, freq = d$freq, min.freq = 1, max.words=200, random.order=FALSE, rot.per=0.25, colors=brewer.pal(8, "Dark2"))

```

## Frequency plot of projects
```{r}
tracker %>%
  group_by(Project) %>% 
  count() %>% 
  ggplot(aes(x=reorder(Project,-n), y=n, fill=-n))+
  geom_bar(stat="identity")+
  xlab("Project")+
  ylab("Count")+
  coord_flip()+
  theme_bw()+
  theme(panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    legend.position = "None")
```



# Auxiliary info 
```{r}
# named list for site codes
sites <- c("TAS"="Tasmania", "SFB"="San Francisco Bay", "KB"="Kane'ohe Bay",  "HAK"="Hakai", "FP"="Fort Pierce",  "FHL"="Friday Harbor", "EW"="Edgewater",  "CB"="Carrie Bow Cay",  "BT"="Bocas del Toro",  "TGC"="Texas Gulf Coast")

sites
```


# "Tetris"" Plot
```{r}
selected_project <- "Benthic Photoquadrats"

tracker %>% ggplot(aes(Year, Site))+
  geom_tile(fill="#97c8eb", color="black", size=0.5)+
  coord_equal()+
  theme_bw() +  # change theme simple with no axis or tick marks
    theme(#panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          plot.title = element_text(hjust = 0.5),
          panel.grid.minor = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.position="bottom", # position of legend or none
          legend.direction="horizontal", # orientation of legend
          legend.title= element_blank(), # no title for legend
          legend.key.size = unit(1.0, "cm"),
          axis.text.x=element_text(size=10),
          axis.text.y=element_text(size=10)
          )

```



```{r, fig.height=12, fig.width=14}

trackerFake <- read_csv(here("data/data-status-tracker-fake.csv"))



# make a data frame with all combinations of Projects, Sites, Years
filled <- trackerFake %>% expand(Site, Year, Project) %>% 
  left_join(trackerFake) %>% mutate(level=case_when(
  L1 == "yes" ~ 1,
  L0 == "yes" ~ 0,
  TRUE ~ NA_real_ 
)) 


filled %>%
  mutate(siteName = sites[Site]) %>% 
  ggplot(aes(Year, fct_reorder(siteName, desc(siteName))))+
        geom_tile(aes(fill=factor(level)), color="black", size=0.1)+
        scale_fill_manual(values=c("#c6eafb","#0d5388"), na.value = "white", breaks=c(0,1), labels=c(" Level 0", " Level 1"))+
        coord_equal()+
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.grid.minor = element_blank(),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              legend.position="right", # position of legend or none
              legend.direction="vertical", # orientation of legend
              legend.title= element_blank(), # no title for legend
              axis.text.x=element_text(size=10, angle=90, vjust=0.5),
              axis.text.y=element_text(size=10)
              ) + facet_wrap(~Project, ncol = 8)
```

```{r}
selected_project <- "Squidpops"

filled %>% 
  mutate(siteName = sites[Site]) %>% 
  filter(Project==selected_project) %>% 
  ggplot(aes(Year, fct_reorder(siteName, desc(siteName))))+
        geom_tile(aes(fill=factor(level)), color="black", size=0.1)+
        scale_fill_manual(values=c("#c6eafb","#0d5388"), na.value = "white", breaks=c(0,1), labels=c(" Level 0", " Level 1"))+
        coord_equal()+
        ggtitle(selected_project)+
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              plot.title = element_text(hjust = 0.5),
              panel.grid.minor = element_blank(),
              axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              legend.position="right", # position of legend or none
              legend.direction="vertical", # orientation of legend
              legend.title= element_blank(), # no title for legend
              axis.text.x=element_text(size=10),
              axis.text.y=element_text(size=10)
              ) 

```


```{r}
filled %>% mutate(siteName = sites[Site])
```

